ODE and Diffusion Models of Tumor Growth

### 2024-03-16 paper.tex
12.25

This is our diffusion equation.

>>>>

Now we will derive the diffusion equations for $\beta$ and $\sigma$. Since we are assuming to be in diffusive equilibrium, we can greatly simplify the expressions.

Recall Fick's second law:

\begin{equation}
    \frac{\partial J}{\partial t} = k \nabla^2J + S(r, t)
\end{equation}

where the $S(r, t)$ term represents the source rate.
We can then formulate our diffusion equations by simplifying this equation. Since we are in the steady state, we know that the rate of change
of concentration is equal to zero (or otherwise we would still have diffusion occuring). So we plug in for the $\nabla^2J$
Laplacian term:

\begin{equation}
    \nabla^2J = \frac{-S(r, t)}{k}
\end{equation}

By assuming spherical symmetry, we can simplify this Laplacian into spherical coordinates:

\begin{equation}
    \nabla^2J = \frac{1}{r^2} \frac{\partial}{\partial r} r^2 \frac{\partial}{\partial r} J
\end{equation}

First, solve for the diffusion equation for $\beta$. As long as we are in the core, by assumption (i), $-S(r, t)$ will be equal to $P$.

\begin{equation}
    \frac{1}{r^2} \frac{\partial}{\partial r} r^2 \frac{\partial}{\partial r} \beta(r, t) = \frac{-P}{k'}
\end{equation}

Then use the step function to make sure that the necrotic core assumption is held:

\begin{equation}
    \frac{1}{r^2} \frac{\partial}{\partial r} r^2 \frac{\partial}{\partial r} \beta(r, t) = \frac{-P}{k'} H(r_1(t) - r)
\end{equation}

This makes the right-hand side of the equation equal to zero when $r_1(t) < r$. This equation now takes the form of (4.3) from Greenspan (1972).

Doing the same process for $\sigma$: First, use assumption (iii) to replace the rate $S(r, t)$ with $a$:

\begin{equation}
    \frac{1}{r^2} \frac{\partial}{\partial r} r^2 \frac{\partial}{\partial r} \sigma(r, t) = \frac{a}{k}
\end{equation}

Then add the appropriate step function conditions:

\begin{equation}
    \frac{1}{r^2} \frac{\partial}{\partial r} r^2 \frac{\partial}{\partial r} \sigma(r, t) = \frac{a}{k} H(r - r_1(t)) H(r_3(t) - r)
\end{equation}

This is equivalent to equation (4.4).

\subsection{Solving the Diffusion Equations}

To solve these partial differential equations, boundary equations need to be established. The boundary conditions have been briefly mentioned at the beginning of this
section. First, $\sigma$ and its partial derivatives need to be continuous, along with $\beta$ and its partial derivatives. It is also necessary for equations
(17) and (18) to be satisfied to guarantee a unique solution. Only one initial condition is necessary to solve the diffusion equation: the inital size of the outer
tumor, $r_3(0)$. It is assumed that at $t = 0$, the tumor is small enough such that $r_1(t) = r_2(t) = 0$.
